### 2011-11-10 ipmu2012/proposal.tex
update

The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.

\subsection{Representation of ill-known valid-time intervals}
Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.

\begin{definition}
A possibilistic valid-time period \emph{PVT} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
\begin{equation}
PVT = \left[X,\ Y \right] 
\end{equation}
X and Y are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
\end{definition}

\subsection{Storage of valid-time intervals}

\subsection{Querying ill-known valid-time intervals}

\subsubsection{Query specification}

\subsubsection{Query evaluation}

\subsubsection{Ranking}