### 2011-11-30 ipmu2012/conclusions.tex
Update for the proposal and the conclusions section.

We presented a valid-time model to represent and querying ill-known temporal intervals. The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes. The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.  As future work, the time interval in the query specification would also be ill-known.