### 2011-12-13 ipmu2012/conclusions.tex
Also the conclusions...

We presented a Valid-time model to represent and query ill-known temporal intervals. The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represents and handle querying temporal uncertainty in the database. The final ranking step provides a usefull result record set. The framework models the Allen relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints. As future work, allowing the time interval in the query specification to be ill-known is considered.