### 2012-01-12 ipmu2012/conclusions.tex
The very freaking final but one stupid sentence too long version of the paper.

We presented a valid-time model to represent and query ill-known temporal intervals. The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represent and handle querying of temporal uncertainty in a database. 
%The final ranking step provides a useful result record set. 
It is possible to model the Allen's relations and also more complex relations. As future work, allowing the time interval in the query specification to be ill-known and new ways for aggregation and ranking are considered.