### 2012-02-25 trunk/book/sources/temporalDatabases.tex
update with the simplification of the text.

%
% Temporal databases
%
A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}. In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained. In subsections \ref{subsubsec:primary-key} and \ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed. Finally, subsection \ref{Comm-temp} presents an overview of some commercial temporal database systems.

\subsection{\label{subsec:td-general-concepts}Basic Concepts and Properties}
A database models some part of reality. As mentioned in the introduction, the part of reality a temporal database tries to model, contains some temporal aspects. For example, in this part of reality, some concepts or objects could have time-related or time-variant properties. The modelling of these temporal aspects has to be handled specifically in order for the database to maintain a consistent model of reality.

Thus, a temporal database will contain temporal values, i.e. values representing (indications of) time. Temporal values in a temporal database can be classified into the following types based on their interpretation and modelling purpose. The definitions and explanations of these types can be found in \cite{Dyreson1994} and \cite{Nascimento95decisiontime} and more information can be found in \cite{Jensen:1991:IIM:627283.627484}, \cite{Snodgrass:1984:TQL:588011.588041} and \cite{Nascimento95decisiontime}.

\begin{svgraybox}
\vspace{-10pt}
\begin{definition}\textbf{Valid Time}\\
The \emph{\textbf{valid time}} (VT) of a fact is the time when the fact is true in the modeled reality.
\end{definition}

\begin{definition}\textbf{Transaction Time}\\
A database fact is stored in a database at some point in time, and after it is stored, it is current until logically deleted. The \emph{\textbf{transaction time}} (TT) of a database fact is the time when the fact is current in the database and may be retrieved.
\end{definition}

\begin{definition}\textbf{Decision Time}\\
\emph{\textbf{Decision time}} (DT) denotes the time when an event was decided to happen.
\end{definition}

\begin{definition}\textbf{User-defined Time}\\
\emph{\textbf{User-defined time}} (UDT) is an uninterpreted attribute domain of date and time.
\end{definition}
\vspace{-10pt}
\end{svgraybox}

\emph{Valid times} are usually supplied by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}. Temporal values of type UDT are not given any special interpretation and have thus no special query language support \cite{Dyreson1994}.

A \emph{temporal database} can now formally be defined as follows \cite{Dyreson1994}.

\begin{svgraybox}
\vspace{-10pt}
\begin{definition}\textbf{Temporal Database}\\
A \emph{\textbf{temporal database}} supports some aspect of time, not counting user-defined time.
\end{definition}
\vspace{-10pt}
\end{svgraybox}

In a relational temporal database, temporal values will of course be in the tuples of the extentions of temporal relations:

\begin{svgraybox}
\vspace{-10pt}
\begin{definition}\textbf{Valid-time Relation}\\
A \emph{\textbf{valid-time relation}} is a relation with exactly one system supported valid time \cite{Dyreson1994}.
\end{definition}

\begin{definition}\textbf{Transaction-time Relation}\\
A \emph{\textbf{transaction-time relation}} is a relation with exactly one system supported transaction time \cite{Dyreson1994}.
\end{definition}
\vspace{-10pt}
\end{svgraybox}

A \emph{valid-time}, respectively \emph{transaction-time} \emph{relational database} is now defined as containing one or more valid-time, respectively transaction-time relations \cite{Dyreson1994}. Next to this, \emph{bitemporal} relational databases contain both valid time and transaction time and tritemporal databases contain valid time, transaction time and decision time.

%Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.

A very extensive list of the most well-known temporal database models can be found in ~\cite{Yu1998}. As it is of course necessary to define a consistent way to query the temporal data, there are several proposals concerned with query languages and query language adaptations for temporal databases like ~\cite{TSQL} and ~\cite{Snodgrass98}.

%A \emph{temporal database}~\cite{etzion1998} is a database that manages some aspects of time in its schema~\cite{Dyreson1994}. The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior. A very extensive list with the most well-known models in temporal databases can be found in ~\cite{Yu1998}. Nevertheless, it is necessary to define some consistent way to query the temporal data. There are several languages for querying temporal databases like TSQL~\cite{TSQL}. In~\cite{Snodgrass98} a proposal to add temporal support to the standard SQL is given.

%\begin{svgraybox}
%The temporal notions in temporal databases can be classified into the following types based on their interpretation and modelling purpose. User-defined time has no interpretation, but the other types do:\\

%	\textbf{Transaction time} \emph{TT}~\cite{Jensen:1991:IIM:627283.627484}: The time when the fact is stored in the database.\\

%	\textbf{Valid time} \emph{VT}~\cite{Snodgrass:1984:TQL:588011.588041}: The time when the fact is true in the modelled reality.\\

%	\textbf{Decision time} \emph{DT}~\cite{Nascimento95}: The time when an event was decided to happen. \\
%\end{svgraybox}

In the following sections, one of the main points of attention concerning relational temporal databases is explained and discussed.

\subsection{\label{subsubsec:primary-key}Primary Keys in Temporal Relations}
When designing a relation, a subset of the relation's attribute set is ususally chosen as primary key. The values of a record for these attributes will then uniquely determine this record and no two records may have the same values for every attribute in this primary key. In temporal relations, different records can exist with the same values for every attribute except the attributes describing the temporal aspects of the objects or concepts represented by the records (the temporal attributes). These different records represent different temporal versions of the same real objects or concepts. To allow the existence of such records, the most common solution is to extend the original primary key with the temporal attributes.
%One of the first tasks in the design of a database is the choice for the primary key. In a temporal database, the most common solution for the choice of the primary key is to extend the primary key with the temporal value. Nevertheless, when dealing with uncertainty in the temporal values, the solution is usually to create a version identifier otherwise the primary key have uncertainty which should be avoided.

The following example illustrates this primary key issue in a valid-time relation.

\begin{example}
\label{ex:pk}
Consider the example database visualized in table \ref{table:example-database}. It contains data representing employees in a company, and two temporal values which represent a valid-time interval. Consider now that we want to hire Sarah again. That is not possible because of the primary key (ID) does not allow to insert again the row Sarah but with a different start and end time. In some models this is resolved by adding to the primary key both values Start and End years. The resulting database allows to insert a new row where Sarah the year 2010. But this modification also allows to insert spurious values e.g. inconsistent time periods (see last row in table \ref{table:example-database-with-new-pk}).

%\vspace{-10pt}

\begin{table}
\centering
\caption{Example database}
\begin{tabular}{c c c c c c }
\hline
\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
1 & Peter & 24 & John &  2010 & - \\
2 & Maria & 34 & John & 2001 & - \\
3 & John & 52 & - &  1999 & - \\
4 & Sarah & 29 & Maria &  2005 & 2009 \\
\hline 
\end{tabular}
\label{table:example-database}

%\vspace{10pt}

\end{table}

%\vspace{-25pt}

%\vspace{-10pt}

\begin{table}
\centering
\caption{Example database with new primary key and some spurious values}
\begin{tabular}{c c c c c c }
\hline
\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
1 & Peter & 24 & John &  2010 & - \\
2 & Maria & 34 & John & 2001 & - \\
3 & John & 52 & - &  1999 & - \\
4 & Sarah & 29 & Maria &  2005 & 2009 \\
4 & Sarah & 29 & Maria &  2010 & - \\
4 & Sarah & 29 & Maria &  2007 & 2008 \\
\hline 
\end{tabular}
\label{table:example-database-with-new-pk}

%\vspace{10pt}

\end{table}

%\vspace{-25pt}

\end{example}

\subsection{\label{subsubsec:consistency}Consistency on temporal databases}
The consistence mechanism in a temporal database is usually re-defined. The main problem is the insertion of spurious values in the database, as shown in example \ref{ex:pk}, in table \ref{table:example-database-with-new-pk}. The most usual solution is to re-define the DML (\emph{Data Manipulation Language}). e.g. the \emph{update} sentence is redefined as two sentences an update and a create sentence as illustrated in the following example.

\begin{example}
Consider the database in table \ref{table:example-database}. The following \emph{update} sentence:

\begin{verbatim}
Update Employees set StartYear = 2010 where ID=3;
\end{verbatim}

Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. The update sentence showed above is translated into:

\begin{verbatim}
Update Employees set EndYear = 2010 where ID=3;
Insert into Employees values (3,John,52,-,2010,-);
\end{verbatim}

%\vspace{-10pt}

\begin{table}
\centering
\caption{Example database after update sentence}
\begin{tabular}{c c c c c c }
\hline
\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
1 & Peter & 24 & John &  2010 & - \\
2 & Maria & 34 & John & 2001 & - \\
3 & John & 52 & - &  1999 & 2010 \\
3 & John & 52 & - &  2010 & - \\
4 & Sarah & 29 & Maria &  2005 & 2009 \\
\hline 
\end{tabular}
\label{table:example-database-update}

%\vspace{10pt}

\end{table}

%\vspace{-25pt}

Table \ref{table:example-database-update} shows the resulting database after the update sentence.

\end{example}

\subsection{\label{Comm-temp}Commercial Temporal database systems}
There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} is a comparative among the most well-known systems. It is also provided a reference for more information.

Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in an OracleDB. Nevertheless TimeDB and Postgree Temporal~\cite{posgree2009} are similar. Both are simple implementations that implement a subset of the Allen's operators and some operations for the creation and manipulation of a temporal attribute (valid, transaction or both times) in a table. Nevertheless, Teradata~\cite{teradata2011} is mainly a business intelligence database designed for data-mining.

The most complete implementation is the Workspace Manager. Unfortunately, none of these systems work with uncertainty or imprecision neither in the data storage nor in the query.

\begin{table}
\centering
\caption{Commercial implementations for temporal database systems. }
\begin{tabular}{c c c c c c }
\hline
\textbf{Name} & \textbf{Time managed} & \textbf{Comments} & \textbf{References}  \\ \hline
Oracle Workspace Manager & VT and TT. & Package for Oracle DB. & \cite{oracle2009}\\
TimeDB & VT and TT. & Interface for Oracle DB. & \cite{timedb2005}\\
Postgree Temporal & VT. & Package for Postgree SQL. & \cite{posgree2009}\\
Teradata & VT and TT. & Used for data-mining. & \cite{teradata2011}\\
\hline 
\end{tabular}
\label{table:commercial-temporal-db}

%\vspace{10pt}

\end{table}

%\subsubsection{Oracle Workspace Manager}
%
%\begin{figure}
%\centering
%\includegraphics[scale=0.5]{graphs/workspaceTopology.eps}
%\caption{A sample topology in the workspace manager.}
%\label{fig:workspace-topology}
%\end{figure}
%
%Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table. The main benefits of this are two:
%
%\begin{itemize}
%\item
%Organization and optimization of data in a hierarchical frames: It is possible to create a frame of time (a workspace) and work only with that partial version of the data. It is provided a mechanism to merge data among different workspaces and to solve conflicts. The organization of data in these workspaces for very large tables result in smaller time access.
%\item
%Valid time as well as transaction time are managed by the system. Each update sentence makes a change in the versioning of a row.
%\end{itemize}
%
%When a table is versioned, the system creates a few tables and views as well as auxiliary structures to allow keep several version of the data, while keeping the primary key defined by the user.
%
%A workspace is a logical group of a set of changes (versions of tuples) and allows consistent access, thus the user always obtain the correct data version. Workspaces are ordered in a hierarchy. The top level is called LIVE workspace (see figure \ref{fig:workspace-topology}).  \\
%
%The system only makes a copy of a tuple if it is changed. In order to access different versions of the tuple, the context must be changed. Also, it is not necessarily to modify the SQL code to access to versioned tables.\\
%%When a table is versioned, it is renamed as table\_LT. Here is stored all the data for the information together metadata for the versioning. An auxiliary table is created with the workspace metadata with the name table\_AUX. A view is created with the name of the original table to allow querying without versioning.