### 2012-03-16 book/sources/proposal.tex
update from the task list.

% Proposal
%
%
The model deals with uncertainty in the valid-time. A valid-time period  \emph{VT} is specified with two points: a starting time point \emph{St} and an ending time point \emph{Et}. Thus, a fuzzy validity period, \emph{FVP} is a valid-time period in which the starting, the ending or both points are not precisely known. More formally we make the following definitions:

\begin{definition}
Valid Time Period: bien
\end{definition}

\subsection{Representation}

\subsubsection{Undelying domain: Julian Day Number}

\subsubsection{Temporal Operators}

\subsection{Data Manipulation Language}

\subsection{Query specification}

\begin{example}

\end{example}