### 2012-04-09 trunk/ipmu2012/conclusions.tex
I think this should be the final version... but how knows..

We presented a valid-time model to represent and query ill-known temporal intervals. The framework tries to correctly represent and handle querying of temporal uncertainty in a database. It is possible to model the Allen's relations and also more complex relations. As future work, new ways for aggregation and ranking are considered.