### 2012-04-10 ipmu2012_2/conclusions.tex
Part of the paper

In this paper, a technique for querying valid-time relations is presented. This technique allows defining a temporal constraint based on a (partially) unknown time interval. The main advantage of this technique is that it tries to correctly represent and handle uncertainty in the temporal constraint, by using ill-known intervals. The technique allows the usage of both the Allen relations and more complex constructions. For future work, allowing the temporal specifications in the database to be (partially) unknown and new methods for aggregation and ranking are considered.

%We presented a valid-time model to represent and querying ill-known temporal intervals. The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes. The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.  As future work, the time interval in the query specification would also be ill-known.