### 2012-04-16 ijcis-pos-model/sources/conclusions.tex
version sent to olga and nicolas.

We presented a valid-time model to represent and handle ill-known temporal intervals. The semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definition Language as well as the querying will be considered.