### 2012-06-12 ijcis-pos-model/sources/abstract.tex
preliminary final modifications

%%%%%%%%%%%%%%%%%%%%%%%%
%
% ABSTRACT
%
%%%%%%%%%%%%%%%%%%%%%%%%
In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature. Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database. Therefore, temporal database models have been proposed to deal with this problem in the literature. Time can be affected by imprecision, vagueness and / or  uncertainty, since existing time measuring devices are inherently imperfect. Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise. However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropiate support in order to accomplish this task. Several proposals for dealing with such imperfections when modelling temporal data exist. Some of these proposals transform the temporal data into a compact representation. There is not a model for managing and handling uncertainty regarding temporal information.
In this work we present a novel model to deal with imprecision in valid-time databases toghether with the data manipulation language, \emph{DML} is defined and implemented.