### 2012-06-13 ijcis-pos-model/sources/conclusions.tex
almost final version

In this work, we presented the first formalization in the literature of a possibilistic valid-time model for a relational databases. First we introduced the representation of the ill-known time intervals. Then, by using the ill-known constraints we have shown how to manage the relationships among time intervals. Then, the valid-time model for relational databases is implemented. Finally, the semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definition Language as well as the querying will be considered.