### 2012-06-20 trunk/ijcis-pos-model/sources/conclusions.tex

In this work, we presented a complete valid-time model to represent and handle ill-known temporal intervals. The paper includes the formal definition of possibilistic valid-time period in order to manage the time and the formal definition of ill-known constraints to define operators and integrity. This is the first formal model in the literature for possibilistic valid-time in relational databases. The semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definition Language as well as the querying will be considered.