### 2013-02-15 fqas2013/paper.tex
The start of the paper.

%%FQAS 2013 Contribution

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\begin{document}

\mainmatter  %start of the contribution

\title{Bipolar Valid-time Querying}

%Short form of title for running heads
\titlerunning{Bipolar Valid-time Querying}

\author{Christophe Billiet\inst{1} \and Jos\'{e} Enrique Pons\inst{2} \and Olga Pons\inst{2} \and Guy De Tr\'{e}\inst{1}}

\authorrunning{Christophe Billiet et al.}

\institute{
Department of Telecommunications and Information Processing, Ghent University,\\
Sint-Pietersnieuwstraat 41, B-9000, Ghent, Belgium\\
Christophe.Billiet@UGent.be, Guy.DeTr\'{e}@UGent.be\\
\and
Department of Computer Science and Artificial Intelligence, University of Granada,\\
C/Periodista Daniel Saucedo Aranda, S/N, E-18071, Granada, Spain\\
jpons@decsai.ugr.es, opc@decsai.ugr.es
}

\maketitle

\begin{abstract}
Here goes the abstract.
\keywords{keyword1, keyword2}
\end{abstract}

\section{Introduction}

\bibliographystyle{splncs}
\bibliography{paper}

\end{document}