### 2013-03-06 eusflat2013/conclusions.tex
introduction, abstract and conclusions

In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information. Besides the differences in the visualization of imperfect time intervals, we obtained that the temporal reasoning in both frameworks is equivalent. It is also important to notice that whereas the ill-known constraint framework deal with the thirteen Allen's relations, the triangular model distinguishes fifteen possible relational zones and therefore, it is necessary to set a correspondence between the Allen's relations in both frameworks.

Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.