### 2013-03-09 eusflat2013/introduction.tex
Some slight changes. I wrote an email about them...

%1. Introduction
Information systems (IS) in general and database systems (DS) in particular have always tried to model parts of reality. To achieve this modelling, such systems contain data representing properties of real-world objects or concepts. As time is an essential aspect of many of these, information systems often contain data representing temporal notions which describe their temporal properties. Such temporal notions usually take the form of either instants (time points)~\cite{Dyreson1994},~\cite{Jensen1998} or time intervals~\cite{Dyreson1994},~\cite{Jensen1998}.
%TODO: more references in the above part?

Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain. Data representing temporal notions may contain such uncertainties too.

Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data. Information or knowledge is usually retrieved from an IS by querying its contents and examining the query results or by visualizing its contents, querying this visualization and examining these query results.
%TODO: include some form of proof or previous work with these statements!

Of course, when temporal information is contained in an IS, it can be used in the querying process. Usually, querying an IS containing data representing temporal notions is done by specifying one or more time indications and requesting information that is related to these notions, where the nature and semantics of these relationships are specifically temporal. Thus, specific temporal relationships are necessary. Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants). The resulting basic Allen relationships are shown in figure \ref{fig:allen-relationships}.

To be able to query IS and DS containing uncertain data representing uncertain time indications, a framework is necessary, able to represent uncertainties in time indications without (much) information loss and able to temporally reason with such uncertain time indications. Although many proposals for such frameworks exist, the work presented in this paper focusses on two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.

The work presented in this paper consists of a comparison of the approaches, used by both frameworks, to representing time intervals subject to uncertainty and to reasoning about the temporal relationships between such intervals and perfectly defined time intervals.

The structure of this paper is as follows: Section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper. Section \ref{sec:ikc} introduces the IKC framework and Section \ref{sec:tm} introduces the TM framework. In Section \ref{sec:proposal}, both frameworks are compared. Finally, Section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

%The concept of the time has been widely studied \cite{Benthem1982}, \cite{Shackle1961}, \cite{Klein1994}. Moreover, humans beings manage temporal indications in an imprecise way \cite{Devos1998}. But when dealing with time in an Information System, some simplifications must to be done. The first thing to do is the discretization of the time line into time points or time intervals. Both approaches have been proved to be equivalent, although due to the nature of the smallest unit of time in a computer, a \emph{chronon}, the discretization of a time point returns a time interval.

%All the possible positions between two time intervals were studied by Allen \cite{Allen1983}, \cite{Allen1985}. As result, the thirteen Allen's Relations were obtained (See Table \ref{tab:allen-relations}) which are illustrated in Figure \ref{fig:allen-relationships}. In order to achieve a more intelligent processing of time, some theoretical frameworks are used to reason with time. First the possibility theory was employed in the reasoning with temporal information \cite{Dubois2003a}. Then, several proposals \cite{Schockaert2008}, \cite{nagypal2003}, \cite{Ohlbach2004}, \cite{Pons2011} studied how to extend the thirteen Allen's relations to the possibilistic case. Also rough set theory \cite{Pawlak1995} has been used to represent and reason about imperfect time intervals \cite{Qiang2009}, \cite{Qiang2010}.

%Although there exists several proposals to represent and visualize imperfect time intervals, in this paper we will focus on two: the ill-known constraint framework \cite{Pons2011} and the triangular model \cite{DeTre2012}. The first one is a theoretical framework that deals with temporal reasoning and the second one is a visual framework that represent imperfect time intervals in a two dimensional space. Both frameworks can be used to represent imperfect time intervals as well as temporal reasoning.

%The structure of this paper is the following. Section  \ref{sec:preliminaries} introduces the Ill-known constraint framework. Section \ref{sec:triangular-model} introduces the triangular model. Section \ref{sec:proposal} analyze both frameworks and shows the correspondences and differences between both frameworks. The section concludes with an example showing that the calculations made in both frameworks are equivalent. Finally, Section \ref{sec:conclusions} presents the main conclusions and the future work.

%\begin{table}[h]
%\centering
%\begin{tabular}{|c|l|}
%\hline
%Name & Implementation \\ \hline 
%$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
%$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
%$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
%$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
%$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
%$I$ meets $J$ & if $e_i = s_j $ \\
%$I$ met by $J$ & if $s_i = e_j $ \\
%$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
%$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
%$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
%$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
%$I$ before $J$ & if $e_i < s_j $ \\
%$I$ after $J$ & if $s_i > e_j $ \\
%\hline
%\end{tabular}
%\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
%\label{tab:allen-relations}
%\end{table}

\begin{figure}[h]
   \centering
   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
   \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
   \label{fig:allen-relationships}
 \end{figure}